---
title: "P8106-HW3-yz4184"
author: "Yunlin Zhou"
output:
  pdf_document:
    toc: yes
    toc_depth: 2
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
header-includes:
- \usepackage{fancyhdr}
- \usepackage{lipsum}
- \pagestyle{fancy}
- \fancyhead[R]{\thepage}
- \fancypagestyle{plain}{\pagestyle{fancy}}
--- 

\newpage


  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r, echo = T, message = FALSE, results='hide', warning=FALSE}
library(caret)
library(glmnet)
library(pROC)
library(pdp)
library(vip)
library(AppliedPredictiveModeling)
library(earth)
library(tidyverse)
library(ggplot2)
library(patchwork)
```

## Data cleaning

```{r, message = FALSE, results='hide', warning=FALSE}
# import data
dat = read.csv("./auto.csv")%>%
  na.omit() %>% 
  mutate(
    cylinders = as.factor(cylinders),
         year = as.factor(year),
         origin = as.factor(origin),
    mpg_cat = factor(mpg_cat, levels = c("low", "high")))
```

```{r}
# divide data into two parts (training and test)
set.seed(1)
rowTrain <- createDataPartition(y = dat$mpg_cat,
                                p = 0.7,
                                list = FALSE)
```

```{r, message = FALSE, results='hide', warning=FALSE}
train_df = dat[rowTrain,]
test_df = dat[-rowTrain,]
```

# (a) Produce some graphical or numerical summaries of the data.

## graphical summaries of continuous variables
```{r warnings = FALSE, fig.height = 4}
theme1 <- transparentTheme(trans = .4)
trellis.par.set(theme1)

featurePlot(x = dat[, 2:5], 
            y = dat$mpg_cat,
            scales = list(x = list(relation = "free"), 
                          y = list(relation = "free")),
            plot = "density", pch = "|", 
            auto.key = list(columns = 2))
```

As shown in the density plots above, we can conclude that the cars with high miles per gallon are tending to have lower weights; larger time to accelerate from 0 to 60 mph; lower engine displacement and lowerhorse power.

## graphical summaries of catagorical variables

```{r warnings = FALSE, fig.height=9}
p_cylinders = dat%>%
  ggplot(aes(x = dat[,1], fill = mpg_cat)) + 
  geom_bar(stat = "count", 
           position = position_dodge(),
           alpha = 0.6)+ 
  labs(
    x = "Number of cylinders",
    y = "The count of cars of different cylinder number"
  )

p_year = dat%>%
  ggplot(aes(x = dat[,6], fill = mpg_cat)) + 
  geom_bar(stat = "count", 
           position = position_dodge(),
           alpha = 0.6)+ 
  labs(
    x = "Model year",
    y = "The count of cars in different Model year"
  )

p_origin = dat%>%
  ggplot(aes(x = dat[,7], fill = mpg_cat)) + 
  geom_bar(stat = "count", 
           position = position_dodge(),
           alpha = 0.6)+ 
  labs(
    x = "Origin of car",
    y = "The count of cars from different origins"
  )

grid.arrange(p_cylinders, p_year,p_origin, nrow = 3)
```

As we can see from the plot above: 4 cylinders car are tending to have the high miles per gallon; as the time went by, the cars are tending to have high miles per gallon; Many American cars have low miles per gallon.

